---
title: "chapter 2"
author: "Warren Kretzschmar"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 2 problems

### Problem 2M3

```{r}
# Pr(land|earth) = 0.3
# Pr(land|mars) = 1
# Pr(land) = (Pr(land |earth) * Pr(earth) + Pr(land|mars) * Pr(mars)) / 2
# Pr(earth) = 0.5
# Pr(mars) = 0.5
# Pr(earth|land) = Pr(land|earth) * Pr(earth) / Pr(land)
like_e = dbinom(1, 1, 0.3)
like_e
like_m = 1
prior_e = 0.5
prior_m = 0.5
ave_prob = like_e * prior_e + like_m * prior_m
ave_prob
like_e
ave_prob
posterior = like_e * prior_e / ave_prob
post_m = like_m * prior_m / ave_prob
posterior + post_m 
posterior
```

### 2M4
```
Card 1 side 1 = B
Card 1 side 2 = B
Card 2 side 1 = B
```
Ways to produce B/B:
```
Card 1 = 1 * 2 = 2
Card 2 = 1 * 1 = 1
Card 3 = 1 * 0 = 0

Posterior = 2 / (1 + 2) = 2/3
```
### 2M5
```
Card 1 = 2 * 2 = 4
Card 2 = 1 * 1 = 1
Card 3 = 1 * 0 = 0

Posterior = 4 / 5
```
### 2M6
```
Card 1 = 1 * 2 = 2
Card 2 = 2 * 1 = 2
Card 3 = 3 * 0 = 0

Posterior = 2 / 4
```
### 2M7
```
Card 1,2 = 1 * 2 * 1 * 1 = 2
Card 1,3 = 1 * 2 * 1 * 2 = 4
Card 2,1 = 1 * 1 * 1 * 0 = 0
Card 2,3 = 1 * 1 * 1 * 2 = 2
Card 3,* = 0

Posterior = (2 + 4) / 8 = 0.75
```

### 2H1
```
PriorA = PriorB =0.5
Pr(Twins) = 0.1*0.5 + 0.2 + 0.5 = 0.15
Pr(A|twins) = Pr(Twins|A)Pr(A)/Pr(Twins)
```
```{r}
pr_a_g_t = 0.1 * 0.5 / 0.15
pr_a_g_t
pr_b_g_t = 0.2 * 0.5 / 0.15
pr_b_g_t
```

Next birth:
Multiply the likelihoods times their priors, which are the species posteriors
after having seen one birth of twins:

`Pr(Twins) = Pr(Twins|A) Pr(A) + Pr(Twins|B) Pr(B)`
```{r}
post_t = 0.1 * pr_a_g_t + 0.2 * pr_b_g_t
post_t
```

Let's do it the other way round. No need for prior as both species have 
equal prior:

```
Panda A T,T 0.1 * 0.1 = 0.01
Panda A T,S 0.1 * 0.9 = 0.09
Panda B T,T 0.2 * 0.2 = 0.04
Panda B T,S 0.2 * 0.8 = 0.16

Pr(Twins on second birth)= (0.01 + 0.04) / (0.01 + 0.09 + 0.04 + 0.16)
```

```{r}
(0.01 + 0.04) / (0.01 + 0.09 + 0.04 + 0.16)
```

### 2H2

This is the first posterior from the first answer to question 2H1:

```
Pr(A|twins) = Pr(Twins|A)Pr(A)/Pr(Twins)
```
```{r}
pr_a_g_t = 0.1 * 0.5 / 0.15
pr_a_g_t
```
### 2H3

```
Pr(A|Singleton) = Pr(Singleton|A) * Pr_a_g_t(A) / Pr(Singleton)
Pr(Singleton) = Pr(Singleton|A) * Pr_a_g_t(A) + Pr(Singleton|B) * Pr_b_g_t(B)
```
```{r}
pr_a_g_ts = pr_a_g_t * 0.9 / (pr_a_g_t * 0.9 + pr_b_g_t * 0.8)
pr_a_g_ts
```

### 2H4
```
Pr(+A|A) = 0.8
Pr(+A|B) = 1-0.65 = 0.35
Pr(+B|A) = 1-0.8 = 0.2
Pr(+B|B) = 0.65

Pr(A|+A) = Pr(+A|A) * Pr(A) / Pr(+A)
```
```{r}
pr_a_g_pa = 0.8 * 0.5 / (0.8*0.5 + 0.35 * 0.5)
pr_a_g_pa
```

Redefine Pr(A) as posterior Pr(A|T)

```
Pr(A|+A) = Pr(+A|A) * Pr(A|T) / Pr(+A)
         = Pr(+A|A) * Pr(A|T) / (Pr(+A|A) * Pr(A|T) + Pr(+A|B) * Pr(B|T))
```
```{r}
pr_a_g_tpA =  0.8 * pr_a_g_t / (0.8 * pr_a_g_t  + 0.35 * pr_b_g_t)
pr_a_g_t
pr_a_g_tpA
```


--- This does not work ---
```
Pr(A,+A,T) = Pr(A)Pr(+A,T|A)
Pr(+A,T) = Pr(+A|T)Pr(T)
Pr(+A,T|A) = Pr(+A|T,A)Pr(T|A)
Pr(A|+A,T) = Pr(+A,T|A)/Pr(T|A)
```

Using chain rule:
```
Pr(A,+A,T) = Pr(A)Pr(T|A)Pr(+A|T,A)
           = Pr(T)Pr(A|T)Pr(+A|T,A)
           = Pr(+A)Pr(A|+A)Pr(T|A,+A)
```


--- End of this does not work ---
